---
title: "Final project - Part 2 Continuation"
author: "Joel Mercado-Diaz"
date: "March 17, 2016"
output: html_document
---

**This is a continuation of Part 2 of my final project. To see the first portion of this part, refer to the file Final_project_Mercado-Diaz.ipynb.** 

In this portion of my project, I need to count and store the number of instances each country appears in the dataset. I will also need to create a new table that will show for each country the number of instances it appears in the dataset, and their corresponding lats and lons. Finally I will generate a map that will be saved into a pdf file for later use.

## Step 2. Count the number of instances

Set working directory (set your working directory here!)
```{r}
setwd("../final_project/")
```

Open packages needed for the rest of the exercise
```{r message=FALSE}
library(ggmap)
library(dplyr)
library(ggplot2)
```

Read table and create column labels
```{r}
df <- read.csv("geo_table.csv", header = FALSE)
colnames(df) <- c("id", "country")
```

Use some dplyr magict to create dataframe that group countries by name in one columns and the number of times they appear in the dataset in another column. Then, add new names columns 
```{r}
count_df <- df %>% group_by(country) %>% summarise(n())
colnames(count_df) <- c("country", "count")
count_df
```

## Step 3. Create coordinates for all countries
I am going to use function _geocode_ from the package _ggmap_ which uses a string with the name of a locality and generates lat/lon coordinates for the centroid of this locality.I will add the lat lons as new columns to the dataframe.

```{r message=FALSE}
lats <- c() 
lons <- c()
for(country in count_df$country){
  coords <- geocode(country)
  lats <- c(lats, coords$lat)
  lons <- c(lons, coords$lon)
}
count_df["lat"] <- lats
count_df["lon"] <- lons
count_df
```

## Step 4. Generate a map of points that are proportional to the number of times the country appears in the dataset.
Recall that the count for each country is equivalent to the number of DNA sequences that belong to material collected in this country. I will make use of _ggplot2_ to streamline the process. The first part will create aa base map. Then I use ggplot to impose points over the map. Notice that size=count indicates that size of the points should be proportional to the number for each country that shows up in the column "count" within count_df. At the end I use the function ggsave to generate the pdf.
```{r message=FALSE}
world_map <- borders("world", colour = "gray50", fill = "gray50")
p <- ggplot(count_df, aes(x = lon, y = lat, size = count)) + world_map + geom_point() + ggsave("geo_map.pdf", width = 14 , height = 9)

p
```

Check that the map was created in the directory. Done!
